Pleiotropy of expression quantitative trait loci increases with the number of regulated genes, active cell types and tissues and bound transcription factors

### 2023-02-18 ms/ms08_figures.tex
MS

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% Fig 1: Comorbidity matrix
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\begin{figure}[!tbp]
    \centering
%
    \includegraphics[width=\textwidth]{\floatRelativePath/plthst_perc_tophits_eqtl.py/hist_perc_tophits_eqtl_excl_mhc.png}
%
    \caption{Histogram showing the percentage of GWAS with the percentage of their tophits colocalized with eQTLs.
    I have excluded tophits in in the MHC locus in chromosome 6 between 25,000,000 and 35,000,000,
    because this locus is not included in the colocalization pipeline.}
    \label{fig:hist_perc_tophits_eqtl_excl_mhc}
%
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% Fig 1: Comorbidity matrix
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\begin{figure}[!tbp]
    \centering
%
    \includegraphics[width=\textwidth]{\floatRelativePath/plthtmp_disease_comorbidity_matrix.py/corr_inkscape.png}
%
    \caption{Trait distance based on the Spearman correlation of colocalized eQTL beta values for different genes and samples.}
    \label{fig:gwas_distance}
%
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% Fig 2: Length histogram of regions
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% \begin{figure}[!tbp]
% \centering
% %
% \begin{subfigure}[]{.33\textwidth}
% \textbf{a}
% \\
% \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_pleiotropic_regions.py/regions_100000_length_hist.png}
% \end{subfigure}
% %
% \begin{subfigure}[]{.33\textwidth}
% \textbf{b}
% \\
% \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_pleiotropic_regions_cumsum.py/pltbar_regions_cumsum.png}
% \end{subfigure}
% %
% \caption{Analysis of pleiotropic regions. \textbf{a}, Length distribution of pleiotropic regions. \textbf{b}, Cumulative sum ordered inversely with the number of GWAS classes.} \label{fig:pleiotropy_region_distribution}
% \end{figure}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% Fig 3: Histograms variants vs GWAS, eGene and eTissues
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\begin{figure}[!tbp]
    \centering
%
    \begin{subfigure}[]{.32\textwidth}
        \textbf{a}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/plthst_gwas_egene_etissue.py/hist_gwas.png}
    \end{subfigure}
%
    \begin{subfigure}[]{.32\textwidth}
        \textbf{b}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/plthst_gwas_egene_etissue.py/hist_egene.png}
    \end{subfigure}
%
    \begin{subfigure}[]{.32\textwidth}
        \textbf{c}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/plthst_gwas_egene_etissue.py/hist_etissue.png}
    \end{subfigure}
%
    \caption{Percentage of colocalized eQTL/GWAS variants with different number of GWAS classes (\textbf{a}), eQTL genes
        (\textbf{b}) and eQTL tissues (\textbf{c}).} \label{fig:hist_gwas_egene_etissue}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% Fig 4: Correlation between GWAS, genes and tissue counts
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\begin{figure}[!tbp]
    \centering
%
    \includegraphics[width=0.7\textwidth]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_egene_etissue_corr.png}
%
    \caption{Spearman correlation between the counts of GWAS classes, eQTL genes and eQTL tissues.}
    \label{fig:correlation_counts}
%
\end{figure}

@@

\begin{figure}[!tbp]
    \centering
%
    \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/vep.png}
%
    \caption{Odds ratio of the variants annotated with given variant effect
    predictor (VEP) consequence and given GWAS class count compared with non
    pleiotropic variants with GWAS class count one. The significance is given for the
    q-value with False discovery rate 5\%.}
    \label{fig:vep_consequence}
%
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% Fig 6: Violin plots. eGenes, eTissues and gwas per variants
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\begin{figure}[!tbp]
    \centering
%
    \begin{subfigure}[]{.45\textwidth}
        \textbf{a}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_etissue_y_egene.py/plt.png}
    \end{subfigure}
%
    \begin{subfigure}[]{.45\textwidth}
        \textbf{b}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_y_etissue.py/plt.png}
    \end{subfigure}
%
%\begin{subfigure}[]{.32\textwidth}
%\textbf{c}
%\\
%\includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_etissue_y_gwas.py/plt.png}
%\end{subfigure}
%
    \caption{Mean number of genes in eQTL-tissue pairs (\textbf{a}) and mean number of tissues in eQTL-gene pairs (\textbf{b}).} \label{fig:gwas_egene_etisue_per_variant}
%
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% Fig 7: TF count per GWAS class count
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[!tbp]
    \centering
%
    \begin{subfigure}[]{.45\textwidth}
        \textbf{a}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_rsid_y_remapnr.py/bxplt_remaptf_per_rsid_flank_10.png}
    \end{subfigure}
%
    \begin{subfigure}[]{.45\textwidth}
        \textbf{b}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_pleiotropy_y_remapcrm.py/remapcrm_flank10.png}
    \end{subfigure}
%
    \caption{Analysis of transcription factor binding.
    \textbf{a}, Binding count of transcription factors in the region surrounding pleiotropic eQTLs with a radius of 50 bp as a function of the GWAS class count.
    \textbf{b}, Odds ratio of eQTLs annotated as belonging to a cis regulatory modules vs. non-annotated.} \label{fig:freq_tf_per_variant}
%
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% Fig 9: graphical conclusions
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\begin{figure}[!tbp]
    \centering
%
    \begin{subfigure}[]{0.95\textwidth}
        \textbf{a}
        \\
        \includegraphics[width=\textwidth]{fig/model1.png}
    \end{subfigure}
%
    \begin{subfigure}[]{0.95\textwidth}
        \textbf{b}
        \\
        \includegraphics[width=\textwidth]{fig/model2.png}
    \end{subfigure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 1: Histogram of percentage of explained of GWAS
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[h!]
    \centering
    %
    \includegraphics[width=\textwidth]{\floatRelativePath/plthst_perc_tophits_eqtl.py/hist_perc_tophits_eqtl_excl_mhc.png}
    %
    \caption{}
    %
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 2: Trait distance matrix
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[h!]
    %
    \centering
    %
    \includegraphics[width=\textwidth]{\floatRelativePath/plthtmp_disease_comorbidity_matrix.py/corr_inkscape.png}
    %
    \caption{}
    %
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 2: Length histogram of regions
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% \begin{figure}[!tbp]
% \centering
% %
% \begin{subfigure}[]{.33\textwidth}
% \textbf{a}
% \\
% \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_pleiotropic_regions.py/regions_100000_length_hist.png}
% \end{subfigure}
% %
% \begin{subfigure}[]{.33\textwidth}
% \textbf{b}
% \\
% \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_pleiotropic_regions_cumsum.py/pltbar_regions_cumsum.png}
% \end{subfigure}
% %
% \caption{Analysis of pleiotropic regions. \textbf{a}, Length distribution of pleiotropic regions. \textbf{b}, Cumulative sum ordered inversely with the number of GWAS classes.} \label{fig:pleiotropy_region_distribution}
% \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 3: Histograms of variants vs GWAS, eQTL genes and eQTL biosamples
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[h!]
    %
    \centering
    %
    \begin{subfigure}[]{.32\textwidth}
        \textbf{a}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/plthst_gwas_egene_etissue.py/hist_gwas.png}
    \end{subfigure}
    %
    \begin{subfigure}[]{.32\textwidth}
        \textbf{b}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/plthst_gwas_egene_etissue.py/hist_egene.png}
    \end{subfigure}
    %
    \begin{subfigure}[]{.32\textwidth}
        \textbf{c}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/plthst_gwas_egene_etissue.py/hist_etissue.png}
    \end{subfigure}
    %
    \caption{}
    %
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 4: Spearman correlation between counts of GWAS categories, eQTL genes and eQTL biosamples
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[h!]
    %
    \centering
    %
    \includegraphics[width=\textwidth]{\floatRelativePath/cmpt_count_per_rsid.py/count_per_rsid_gwas_egene_etissue_corr.png}
    %
    \caption{}
    %
\end{figure}

@@

\begin{figure}[h!]
    %
    \centering
    %
    \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_vep_consequence.py/vep.png}
    %
    \caption{}
    %
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 6: beta and pval
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[h!]

    \begin{subfigure}[]{.50\textwidth}
        \textbf{a}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/eqtl_beta.png}
    \end{subfigure}
    %
    \begin{subfigure}[]{.50\textwidth}
        \textbf{b}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/gwas_beta.png}
    \end{subfigure}

    \begin{subfigure}[]{.50\textwidth}
        \textbf{c}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/eqtl_neglog10pval.png}
    \end{subfigure}
    %
    \begin{subfigure}[]{.50\textwidth}
        \textbf{d}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_gwas_cat_y_beta_neglog10pval.py/gwas_neglog10pval.png}
    \end{subfigure}
    %
    \caption{}
    %

\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 7: Plots of the number of genes and biosample per variant-biosample and variant-gene, respectively
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[h!]
    %
    \centering
    %
    \begin{subfigure}[]{.45\textwidth}
        \textbf{a}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_etissue_y_egene.py/plt.png}
        %
    \end{subfigure}
    %
    \begin{subfigure}[]{.45\textwidth}
        \textbf{b}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_egene_y_etissue.py/plt.png}
    \end{subfigure}
    %
    \caption{} \label{fig:gwas_egene_etisue_per_variant}
    %
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Fig 8: TF count per GWAS class count
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{figure}[h!]
    \centering
    %
    \begin{subfigure}[]{.48\textwidth}
        \textbf{a}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/pltbox_x_per_rsid_y_remapnr.py/bxplt_remaptf_per_rsid_flank_10.png}
    \end{subfigure}
    %
    \begin{subfigure}[]{.48\textwidth}
        \textbf{b}
        \\
        \includegraphics[width=\textwidth]{\floatRelativePath/pltbar_x_per_variant_pleiotropy_y_remapcrm.py/remapcrm_flank10.png}
    \end{subfigure}
    %
    \caption{}
    %
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%
%% Fig 9: graphical conclusions
%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
\begin{figure}[h!]
    \centering
%
    \begin{subfigure}[]{0.95\textwidth}
        \textbf{a}
        \\
        \includegraphics[width=\textwidth]{fig/model1.png}
    \end{subfigure}
%
    \begin{subfigure}[]{0.95\textwidth}
        \textbf{b}
        \\
        \includegraphics[width=\textwidth]{fig/model2.png}
    \end{subfigure}
